---
title: "DATA 621 HW 3 Crime"
author: "Jian Quan Chen"
date: "2024-03-22"
output: html_document
---

# 1. Data Exploration

```{r warning=FALSE}
library(tidyverse)
library(psych)
library(corrplot)
```

```{r}
df <- read.csv("https://raw.githubusercontent.com/LeJQC/DATA-621-Group-2/main/HW3/crime-training-data_modified.csv")
```

There are 466 observations and 13 variables in the training dataset. Of the 13 variables, 12 are predictor variables and 1 is the target variable.

```{r}
glimpse(df)
```

The variables are:

-   `zn`: proportion of residential land zoned for large lots (over 25000 square feet)
-   `indus`: proportion of non-retail business acres per suburb
-   `chas`: a dummy var. for whether the suburb borders the Charles River (1) or not (0)
-   `nox`: nitrogen oxides concentration (parts per 10 million)
-   `rm`: average number of rooms per dwelling
-   `age`: proportion of owner-occupied units built prior to 1940
-   `dis`: weighted mean of distances to five Boston employment centers
-   `rad`: index of accessibility to radial highways
-   `tax`: full-value property-tax rate per \$10,000
-   `ptratio`: pupil-teacher ratio by town
-   `lstat`: lower status of the population (percent)
-   `medv`: median value of owner-occupied homes in \$1000s
-   `target`: whether the crime rate is above the median crime rate (1) or not (0)

### Summary Table
```{r}
summary_table <- describe(df)

print(round(summary_table,2))
```

### Distrubtion Plot
```{r}
df_long <- df %>%
  pivot_longer(
    cols = everything(), 
    names_to = "variable",
    values_to = "value"
  )

df_long %>%
  ggplot(aes(value)) + 
  geom_density(fill = "blue") + 
  facet_wrap(~variable, scales ="free", ncol = 4) +
  labs(x = element_blank(), y = element_blank())
```

### Correlation Plot
```{r}
df %>% 
  cor(use = "complete.obs") %>%
  corrplot(method = "color", tl.col = "black")
```

# 2. Data Preparation


























