---
title: "NYC RealEstate"
author: "Fredrick Jones, Jian Quan Chen, Tilon Bobb"
date: "2024-04-24"
output:
  pdf_document: 
    toc: yes
    toc_depth: '2'
    latex_engine: xelatex
  html_document:
    theme: united
    highlight: tango
    number_sections: false
    toc: yes
    toc_depth: 5
    toc_float: yes
editor_options: 
  markdown: 
    wrap: 72
---

## Loading Required libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(ggplot2)
library(tidyr)  
library(kableExtra)
library(corrplot)
library(skimr)
library(dplyr)
library(Hmisc)
library(reshape2)
library(tidyr)  
library(MASS)
library(treemap)
library(randomForest)
library(lubridate)
library(forecast)
library(caret)
library(readxl)

```

```{r}
#Clear all
rm(list = ls())

options(scipen = 999)
```
## Abstract
This study uses real estate transaction data to investigate the factors influencing property values in New York City. The dataset includes a variety of parameters that were gathered from public records and real estate listings, including property location, kind, size, and sale price. To understand the distribution and interrelationships of the dataset, exploratory data analysis (EDA) is the first systematic stage in the study technique. Data preparation is a step that comes next in order to encode categorical variables, handle missing values, and create new features. Stepwise regression, generalized linear models (GLM), robust regression, and conventional linear regression are all included in the design of regression models. The goal of developing predictive models that offer robustness against outliers while generalizing effectively to new data is achieved through the use of goodness-of-fit measures and diagnostic tests for residual analysis as the basis for model selection.

## Keywords 
NYC Real Estate
Data Analysis
Regression Modeling
Housing Market Trends
Neighborhood Analysis

## Introduction 

The New York real estate market is one of the most dynamic and influential sectors of urban development. With every real estate transaction representing a tangible real estate exchange, the NYC real estate market is a barometer of the city's socioeconomic landscape. In this report, we take an in-depth look at the NYC market through insights from a diverse dataset of New York real estate sales records.

### Background and Motivation

The database under study contains a wealth of data spanning. two decades, providing a detailed chronicle of real estate transactions in the five boroughs of New York - Manhattan, the Bronx, Brooklyn, Queens and Staten Island. Our analysis is based on the foundation created by this dataset, which provides insight into the multifaceted dynamics of NYC real estate sales.

The motivation is to identify the factors that influence NYC real estate prices, so our research is based on a variety of analytical techniques. and methods. From data analysis to regression modeling, we seek to uncover the complex interplay of variables that affect real estate. By examining real estate sales trends, identifying key predictors of sales prices, and assessing the impact of factors such as property size, location, and tax bracket, we aim to shed light on the mechanics behind the NYC real estate world.

As we delve into the depths of the NYC real estate market, our analysis aims to provide stakeholders with actionable insights on investors and from decision makers to real estate developers and potential buyers. By uncovering the drivers of real estate prices and delineating market trends, we aim to provide decision makers with the information they need to navigate the complexities of the real estate ecosystem.

## Literature Review


#### [Michael Gaynor's Project](https://medium.com/@mgaynor228/analysis-of-nyc-property-sales-9af7686aa2ca)

Michael Gaynor conducted a project to explore the NYC real estate market using SQL queries and Tableau visualization techniques. His investigation aimed to answer four main questions:

1. Which of the five boroughs is the most expensive?
2. Which of the five boroughs have the most sales?
3. What type of properties sell the most in each of the 5 boroughs?
4. What property type influences sales?

Gaynor's approach involved understanding the task, prepping the data, analyzing the data using SQL queries and Tableau visualization, and presenting the findings through an interactive dashboard. Through his analysis, Gaynor discovered insights such as the most expensive borough, the borough with the most property sales, and the types of properties that sell the most in each borough.

### Comparison and Evaluation

Gaynor's research utilized the same NYC real estate dataset to address similar questions as our investigation. However, there are significant differences between Gaynor's approach and our own project:

- **Methodology**: Gaynor primarily used SQL queries and Tableau visualization tools for data analysis, while our investigation utilized R programming language. Our approach involved a combination of data preprocessing, exploratory data analysis (EDA), statistical modeling, and visualization techniques implemented in R.

- **Presentation of Findings**: Gaynor's project focused on presenting the results through an interactive dashboard created in Tableau. In contrast, our investigation may present the findings through various formats such as tables, charts, and narratives within the R Markdown document.

- **Data Preprocessing**: While Gaynor mentioned data cleaning and preprocessing, the details of these steps were not extensively discussed. In our investigation, we employed specific techniques such as handling missing values, outlier detection and removal, and data transformation using R packages like dplyr and tidyr.

- **Statistical Modeling**: Our investigation may involve the application of statistical models such as linear regression, generalized linear models, or machine learning algorithms to explore relationships between variables and predict real estate prices. Gaynor's project did not explicitly mention the use of statistical modeling.

### Advantages and Drawbacks

The advantages of Gaynor's approach include:

- Comprehensive analysis of the NYC real estate market using SQL and Tableau.
- Clear presentation of findings through interactive visualization.

However, there may be some drawbacks to Gaynor's approach, such as:

- Reliance on SQL and Tableau tools may limit accessibility for researchers unfamiliar with these technologies.
- Lack of detailed explanation of data cleaning and preprocessing steps.

## Regression Modelling Methodology

### Data Preparation

The first part of our analysis consisted of importing the dataset that included data on property sales in New York City. After we uploaded the dataset, we conducted an exploratory data analysis to understand its organization, factors, and any problems that required attention.
During the exploratory data analysis, we faced one of the first hurdles with the discovery of missing values in multiple columns. To tackle this problem, we methodically pinpointed the variables with incomplete data and assessed the percentage of missing values in each instance. After careful deliberation, we made the choice to eliminate data points containing incomplete information, as they comprised less than 5% of the entire data set. This method enabled us to keep a large part of the data while reducing the effect of missing values on future analyses.
After addressing missing data, we focused on the distribution of numerical variables in the dataset. We noticed that many variables showed noticeable skewness, suggesting possible departures from normal distribution. To tackle this problem, we utilized Tukey's method for identifying and eliminating outliers. Our goal was to enhance the robustness of future analyses by improving the distributional properties of numeric variables through the identification and exclusion of outliers from the dataset.
Additionally, we examined the connections between variables using correlation analysis. This included computing correlation coefficients for pairs of numerical variables in order to evaluate the magnitude and orientation of their relationships. During this examination, we found multiple variables that showed strong positive relationships with each other, along with variables that had weaker or negative relationships. These results offered valuable information on potential factors that could predict real estate prices and guided the choice of variables for inclusion in regression analysis.
In addition, we analysed changes over time by graphing the time-based pattern of property values in New York City. This examination showed a rising pattern in mean sale prices over time, with variations in certain time frames. Through the visualization of time-based trends, we achieved a better comprehension of the fluctuations within the real estate market of New York City, pinpointing potential influences on property price fluctuations.


### Regression Modelling

After completing thorough data preparation and exploratory analysis, we focused on the main objective of our research: using regression modelling to forecast real estate prices in New York City. The method we used involved carefully building linear regression models, using different predictor variables to understand the complex factors influencing real estate prices.
Leading our regression modelling was the incorporation of important predictor variables that were considered to have a significant impact on real estate prices. These variables included a wide range of factors, each providing valuable perspectives on the intricate fabric of the New York City real estate market. Included in these predictors were the quantity of housing units in a property, offering an insight into its size and ability to house residents. The categorization of taxes for properties at various points highlighted their financial status and legal ramifications, revealing insights into the larger economic and legal environments in which these properties function.
Furthermore, the year in which the building was constructed was identified as a crucial factor in predicting real estate values, giving a historical perspective on how they change over time. We aimed to capture temporal trends and identify any seasonal or cyclical patterns that could affect pricing dynamics by including the sale date of properties in our models. Moreover, factors like total area and land area were essential in evaluating the physical size and spatial characteristics of properties, enhancing our comprehension of their inherent value.
By conducting thorough regression analysis, we discovered significant statistical connections between these predictor variables and property prices, revealing the complex interaction of factors that influence pricing decisions in the real estate market of New York City. Nevertheless, even though our models were strong, the adjusted R-squared values suggested that there might be additional variability that was not accounted for, indicating the presence of hidden factors that were not included in our analysis.



## Explorartory Data Analysis

### Loading Data

```{r}
nyc_data <- read.csv("C:/Users/bobbt/OneDrive/Documents/azurework/nyc-property-sales.csv")

head(nyc_data)
```


#### Glipmse of the dataset

```{r}
glimpse(nyc_data)

```

Assessing missing values

```{r}
# Check for missing values
missing_values <- colSums(is.na(nyc_data))

# View columns with missing values
missing_columns <- names(missing_values[missing_values > 0])
print(missing_columns)
```

Drop missing values since there is less than 5% of dataset missing values hence safe to drop all missing values

```{r}
clean_nyc <- na.omit(nyc_data)
str(clean_nyc)
```

All numeric variables are heavily skewed to the right,hence a clear indication of outliers

All distributions exhibit a highly skewed pattern, with a single bar extending vertically at the rightmost end of the x-axis, suggesting the presence of potential outliers with extremely high unit counts.

```{r}
# Numeric variables
numeric_vars <- c("RESIDENTIAL.UNITS", "COMMERCIAL.UNITS", "TOTAL.UNITS", 
                  "LAND.SQUARE.FEET", "GROSS.SQUARE.FEET", "SALE.PRICE")


num_data <-clean_nyc[, numeric_vars] 
for (i in 1:length(names(num_data))){
  print(i)
  hist( num_data[i], main='hist', breaks=20, prob=TRUE)
}


```

```{r}
# Function to remove outliers based on Tukey's method
remove_outliers <- function(data, variable) {
  q1 <- quantile(data[[variable]], 0.25)
  q3 <- quantile(data[[variable]], 0.75)
  iqr <- q3 - q1
  lower_bound <- q1 - 1.5 * iqr
  upper_bound <- q3 + 1.5 * iqr
  filtered_data <- data[data[[variable]] >= lower_bound & data[[variable]] <= upper_bound, ]
  return(filtered_data)
}

# Apply the function to each numeric variable in clean_nyc
for (var in numeric_vars) {
  clean_nyc <- remove_outliers(clean_nyc, var)
}

```

There was a clear improvement of distribution after removal of outliers

Distribution after applying an outlier removal technique, such as the Interquartile Range (IQR) method. The blue bars show a somewhat more balanced distribution, with the most extreme outliers removed, resulting in a narrower range of residential unit counts.

```{r}

num_data <-clean_nyc[, numeric_vars] 
# Create histograms for each numeric variable
hist_plots <- lapply(numeric_vars, function(var) {
  ggplot(data = num_data, aes_string(x = var)) +
    geom_histogram(fill = "skyblue", color = "black", bins = 30) +
    labs(title = paste("Histogram of", var),
         x = var,
         y = "Frequency") +
    theme_minimal()
})

# Output the histograms
for (plot in hist_plots) {
  print(plot)
}


# Drop COMMERCIAL.UNITS variable
clean_nyc <- clean_nyc[, !names(clean_nyc) %in% "COMMERCIAL.UNITS"]

```

Categorical variables distributions

```{r}
library(treemap)

# Categorical variables
categorical_vars <- c("NEIGHBORHOOD", "BUILDING.CLASS.CATEGORY", 
                      "TAX.CLASS.AT.PRESENT", "BUILDING.CLASS.AT.PRESENT", 
                      "TAX.CLASS.AT.TIME.OF.SALE", "BUILDING.CLASS.AT.TIME.OF.SALE")

# Create treemaps for each categorical variable
treemap_plots <- lapply(categorical_vars, function(var) {
  treemap(clean_nyc, index = var, vSize = "SALE.PRICE", title = paste("Treemap of", var))
})

# Output the treemaps
for (plot in treemap_plots) {
  plot
}

```

The darker blue circles indicate a stronger positive correlation, while the lighter blue and red circles represent weaker or negative correlations. For example, there is a strong positive correlation between gross square feet and land square feet, as well as between residential units and total units. However, sale price has a weak or slightly negative correlation with most of the other variables, suggesting that higher sale prices may not necessarily be associated with larger property sizes or more units.

```{r}
print(sum(any(is.na(clean_nyc))))

num_data <- as.data.frame(num_data)

# Drop COMMERCIAL.UNITS variable
num_data <- num_data[, !names(num_data) %in% "COMMERCIAL.UNITS"]


# Remove observations with missing, NaN, and infinite values
clean_data <- num_data[complete.cases(num_data) & !is.infinite(rowSums(num_data)), ]

# Calculate correlation matrix
correlation_matrix <- cor(clean_data)

# Plot correlation matrix
corrplot(correlation_matrix, method = "circle", type = "upper", order = "hclust",
         addCoef.col = "black", tl.cex = 0.7, cl.cex = 0.7)


```

Transform Categorical Variables to Factors

```{r}

cat_vars <- c("BUILDING.CLASS.CATEGORY", "TAX.CLASS.AT.PRESENT", "BUILDING.CLASS.AT.PRESENT", "TAX.CLASS.AT.TIME.OF.SALE")

# Convert categorical variables to factors
for (var in cat_vars) {
  clean_nyc[[var]] <- factor(clean_nyc[[var]])
}

# Verify the transformation
str(clean_nyc[cat_vars])


```

## 2.  ANALYSIS

Over the years, there is a clear upward trend, indicating that the average real estate prices in NYC have been steadily increasing. The line exhibits a consistent upward slope, with prices rising from around \$350,000 in 2005 to over \$450,000 by 2020. Although there are some fluctuations in specific years, the overall trajectory demonstrates a significant increase in real estate prices in NYC over the 15-year period depicted in the graph.

```{r}
# Convert SALE_DATE to Date format
clean_nyc$SALE_DATE <- as.Date(clean_nyc$SALE.DATE)


# Group data by year and calculate average sale price per year
yearly_prices <- clean_nyc %>%
  mutate(year = lubridate::year(SALE_DATE)) %>%
  group_by(year) %>%
  summarise(avg_price = mean(SALE.PRICE))

# Create a line plot of average sale price over time (yearly)
ggplot(yearly_prices, aes(x = year, y = avg_price)) +
  geom_smooth(method = "lm", se = FALSE, color = "blue", linetype = "solid", size = 1) +  #smoother line without confidence intervals
  geom_point(color = "blue", size = 3) +  
  labs(title = "Average Real Estate Prices in NYC",
       subtitle = "Yearly Trend",
       x = "Year",
       y = "Average Sale Price",
       caption = "Data Source: NYC Real Estate Dataset") +
  theme_minimal() +  
  theme(plot.title = element_text(face = "bold", size = 18),
        plot.subtitle = element_text(size = 14),
        plot.caption = element_text(size = 10),
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10))  


```

```{r}
str(clean_nyc)
```

### Key Factors Influencing Real Estate Prices

The linear regression model suggests that various factors significantly influence real estate prices in New York City. Notably, residential units, tax class, year built, sale date, tax class at time of sale, gross square feet, and land square feet all demonstrate statistically significant relationships with sale prices. However, the model's adjusted R-squared value of 0.06164 indicates that only about 6.164% of the variability in sale prices is explained by these factors. Additionally, the residuals' distribution reveals a considerable spread, indicating potential heteroscedasticity or unaccounted-for factors in the model.

```{r}
# regression analysis
lm_model <- lm(SALE.PRICE ~ RESIDENTIAL.UNITS + TAX.CLASS.AT.PRESENT + YEAR.BUILT + SALE_DATE + TAX.CLASS.AT.TIME.OF.SALE + GROSS.SQUARE.FEET + LAND.SQUARE.FEET, data = clean_nyc)
summary(lm_model)
```

The stepwise regression process, with a starting AIC of 29457675, selected a final model with predictors including residential units, tax class, year built, sale date, tax class at the time of sale, gross square feet, and land square feet. This model, fitted using linear regression, reveals statistically significant relationships between these predictors and sale prices, as indicated by the low p-values and the coefficients' significance levels. However, the adjusted R-squared value remains low at 0.06164, suggesting that this model explains only a small portion of the variability in sale prices.

```{r}
# Perform stepwise regression
stepwise_model <- step(lm_model)

# Summary of the stepwise model
summary(stepwise_model)

```

Considering that Normality was not satisfied

The next model fitted using generalized linear regression with a Gaussian family and an identity link function maintains predictors including residential units, tax class, year built, sale date, tax class at the time of sale, gross square feet, and land square feet. The coefficients and their significance remain consistent with the previous models. The null and residual deviances provide additional information on the goodness of fit, with the residual deviance being slightly lower than the null deviance, suggesting some level of model improvement.

```{r}
# Fit GLM with different error distribution and link function
glm_model <- glm(SALE.PRICE ~ RESIDENTIAL.UNITS + TAX.CLASS.AT.PRESENT + YEAR.BUILT + SALE_DATE + TAX.CLASS.AT.TIME.OF.SALE + GROSS.SQUARE.FEET + LAND.SQUARE.FEET, 
                 data = clean_nyc, 
                 family = gaussian(link = "identity"))
summary(glm_model)


```

The final model, fitted using robust linear regression (rlm), estimates the intercept at \$290,788.64. Each additional residential unit increases the sale price by \$33,326.57. For the tax class at present, each category shows significant negative impacts on the sale price, with Tax Class 1B having the largest effect, reducing the price by \$436,466.67. A one-unit increase in the year built is associated with a \$9.05 increase in sale price. Similarly, each day increment in the sale date adds \$14.35 to the sale price. Other variables, such as gross square feet and land square feet, also exhibit significant positive effects on the sale price. The residual standard error, indicating the model's accuracy, is \$428,200.

```{r}
library(MASS)

# Fit robust linear regression model
lm_model_robust <- rlm(SALE.PRICE ~ RESIDENTIAL.UNITS + TAX.CLASS.AT.PRESENT + YEAR.BUILT + SALE_DATE + TAX.CLASS.AT.TIME.OF.SALE + GROSS.SQUARE.FEET + LAND.SQUARE.FEET, 
                       data = clean_nyc)
summary(lm_model_robust)


```
